\chapter{Conclusion}

This project developed and applied a quantitative framework for assessing economic resilience under disruption, relevant to Singapore's national security planning.

Using the Dynamic Inoperability Input-Output Model (DIIM), two disruption scenarios were studied. The COVID-19 pandemic scenario (15~sectors, 2019~IO table) demonstrated how lockdown-induced workforce unavailability cascaded through input-output linkages to produce sector-level inoperability paths and cumulative economic losses. The manpower disruption scenario (15~sectors, 2022~IO table), motivated by the risk of grey zone conflicts triggering foreign worker departures, showed how purely structural disruption ($c^* = 0$) produces different cascading dynamics compared to a combined supply-and-demand shock.

This study's main contribution is the systematic comparison of five simplified methods for identifying key sectors against the DIIM benchmark, evaluated across multiple intervention scales ($k = 3, 5, 7, 10, 12$) using 20\,000 Monte Carlo trials with realistic $q(0)$ perturbations. The results show three ranges:

\begin{itemize}
    \item At \textbf{small $k$} ($k = 3$), all simplified methods agree on the dominant sectors and perform identically, achieving $\bar{R}_m$ of 0.73--0.87.

    \item At \textbf{medium $k$} ($k = 5$--$7$), methods diverge and structural information adds substantial value. PCA $\times\, x_i$ outperforms the Total Output benchmark by up to 27\% in the COVID-19 scenario (close rate 64\% vs 6\%) by identifying structurally critical sectors (e.g., Construction) that size-based rankings miss. BL $\times\, x_i$ provides complementary value at $k = 7$ (close rate 67\%). At $k = 5$, the best simplified method recovers only 82\% of the loss reduction that DIIM achieves on average. However, structural methods close much of the gap relative to the Total Output benchmark.

    \item At \textbf{large $k$} ($k = 12$), simplified methods converge to DIIM performance, with $\bar{R}_m$ of 0.92--0.97 and close rates above 99\%. At this scale, even the simplest method (ranking by Total Output alone) is a reliable substitute for the full simulation.
\end{itemize}

Four robustness analyses support these findings. First, the three-ranges structure is stable across intervention magnitudes of 5--30\%, confirming that relative method rankings depend on \emph{which} sectors are selected, not the magnitude of assistance. Second, PCA robustness testing reveals that the optimal number of principal components is scenario-dependent ($n_{\textit{pcs}} = 2$--$3$ for COVID-19, $n_{\textit{pcs}} = 1$ for Manpower), introducing a hyperparameter that cannot be set without knowledge of the disruption profile. Third, intervention timing analysis shows that the DIIM's advantage over simplified methods is greatest when intervention is immediate; with longer delays, all methods converge as the absolute benefit of intervention diminishes. Fourth, scatter plot diagnostics confirm that PCA $\times\, x_i$'s advantage at $k = 5$ is systematic across the $q(0)$ perturbation space, not driven by outlier trials.

The near-equivalence between Total Output and structurally-weighted methods at most $k$ values shows that a sector's absolute economic scale is the dominant predictor of macro-level loss reduction. Structural analysis adds value only in the medium-$k$ range, where the marginal sectors differ across methods, and only when the economy contains sectors with a significant mismatch between economic size and structural importance.

Simplified methods perform best when the disruption profile aligns with the economic size distribution, a condition that is scenario-dependent and cannot be assumed in advance.

For national security planning, this enables a scale-dependent approach: use simplified structural analysis for peacetime resilience investments and for large-scale crisis response ($k \geq 10$), but run the full DIIM when targeting fewer than 10 sectors during a crisis. The computational cost of the DIIM is modest, and the performance gain at small $k$ is substantial.

The framework is generalisable beyond the two scenarios studied here. Any economy with an available IO table can compute the simplified structural rankings in peacetime and apply the multi-$k$ analysis to determine the appropriate tool for its intervention scale. Future work could extend the approach to finer sector disaggregation, correlated disruptions, and multi-period dynamics.

\newpage
