%=== CHAPTER SIX (6) ===
%=== Monte Carlo Robustness Analysis ===

\chapter{Monte Carlo Robustness Analysis}

The previous chapter showed that cheap methods can approach DIIM at specific $q_0$ points but not universally. This chapter uses Monte Carlo simulation to systematically evaluate the robustness of cheap methods across a range of realistic disruption profiles and intervention scales.

\section{Framework}

\subsection{Motivation}

All cheap methods produce a \emph{fixed} sector ranking---they depend only on the IO table structure ($A$, $x$), not on the disruption profile ($q_0$). The DIIM, by contrast, adapts its ranking to the specific $q_0$. A cheap method performs well when its fixed ranking happens to align with where $q_0$ concentrates; it fails when the disruption hits sectors that the cheap method did not prioritise.

This insight suggests that the effectiveness of a cheap method is predictable from (i)~features of $q_0$ and (ii)~the number of sectors targeted for intervention~$k$. At small~$k$, only the very top sectors matter, while at large~$k$ the specific ranking becomes less consequential because most sectors are covered regardless.

\subsection{Performance Metric}

For each cheap method~$m$ and a given $q_0$, we compute the performance ratio from Equation~\eqref{eq:performance-ratio}. The mean ratio $\bar{R}_m$ across trials is the primary metric: it captures how much of the DIIM's loss reduction a cheap method achieves on average. As a secondary metric, we report the \emph{close rate}---the fraction of trials where $R_m \geq 0.80$, i.e., the cheap method captures at least 80\% of the DIIM's benefit.


\section{Monte Carlo Design}

\subsection{Perturbation of Real $q_0$}

Rather than drawing $q_0$ from a uniform distribution---which would generate many unrealistic disruption profiles---we perturb the actual scenario-specific $q_0$ vector using log-normal noise:

\begin{equation}
    q_{0,i}^{\text{trial}} = q_{0,i}^{\text{actual}} \cdot \exp\!\bigl(\varepsilon_i\bigr), \quad \varepsilon_i \sim \mathcal{N}(0,\; 0.5^2)
    \label{eq:q0-perturbation}
\end{equation}

\noindent with the constraint $q_{0,i}^{\text{trial}} \in [10^{-6},\; 1]$. This preserves the realistic structure of the disruption (sectors that were heavily impacted remain relatively more impacted) while introducing meaningful variation. The log-normal noise ensures that the perturbation is multiplicative---a sector with $q_0 = 0.01$ can plausibly move to $0.005$ or $0.02$, but not jump to $0.5$.

\subsection{Simulation Procedure}

For each scenario and each value of $k \in \{3, 5, 7, 10, 12\}$, we draw 2\,000 perturbed $q_0$ vectors. For each trial:

\begin{enumerate}
    \item Perturb $q_0^{\text{actual}}$ using Equation~\eqref{eq:q0-perturbation}.
    \item Run the DIIM baseline (no intervention) and compute total economic loss.
    \item Run the DIIM with intervention in the DIIM's top-$k$ sectors (10\% $q_0$ reduction) and compute $\Delta EL_{\text{DIIM}}$.
    \item For each cheap method~$m$, run the DIIM with intervention in method~$m$'s top-$k$ sectors and compute $\Delta EL_m$.
    \item Compute $R_m = \Delta EL_m / \Delta EL_{\text{DIIM}}$ and record whether $R_m \geq 0.80$ (``close'').
\end{enumerate}

This yields 20\,000 simulation trials per scenario (2\,000 $\times$ 5 values of $k$), for a total of 40\,000 trials across both scenarios.


\section{Results: Performance Across $k$}

\subsection{Mean Performance Ratio}

Tables~\ref{tab:covid-ratio-by-k} and~\ref{tab:manpower-ratio-by-k} present the mean performance ratio $\bar{R}_m$ for each cheap method across all values of~$k$.

\begin{table}[H]
\centering
\begin{tabular}{@{}lccccc@{}}
\toprule
\textbf{Method} & $k{=}3$ & $k{=}5$ & $k{=}7$ & $k{=}10$ & $k{=}12$ \\
\midrule
Total Output         & 0.870 & 0.641 & 0.818 & 0.908 & 0.953 \\
PCA $\times$ xi      & 0.870 & \textbf{0.817} & 0.818 & 0.800 & 0.923 \\
PageRank $\times$ xi & 0.870 & 0.641 & 0.703 & 0.908 & 0.953 \\
BL $\times$ xi       & 0.870 & 0.641 & \textbf{0.834} & 0.908 & 0.953 \\
FL $\times$ xi       & 0.870 & 0.641 & 0.818 & 0.908 & 0.923 \\
\bottomrule
\end{tabular}
\caption{Mean performance ratio (cheap / DIIM) by $k$ for COVID-19 (2\,000 trials per $k$). Bold indicates the best cheap method at that $k$.}
\label{tab:covid-ratio-by-k}
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{@{}lccccc@{}}
\toprule
\textbf{Method} & $k{=}3$ & $k{=}5$ & $k{=}7$ & $k{=}10$ & $k{=}12$ \\
\midrule
Total Output         & 0.652 & 0.682 & 0.927 & 0.926 & 0.976 \\
PCA $\times$ xi      & 0.652 & 0.682 & 0.927 & 0.926 & 0.976 \\
PageRank $\times$ xi & 0.652 & 0.682 & 0.749 & 0.926 & 0.976 \\
BL $\times$ xi       & 0.652 & 0.682 & 0.927 & 0.885 & 0.976 \\
FL $\times$ xi       & 0.652 & 0.682 & 0.927 & 0.926 & 0.976 \\
\bottomrule
\end{tabular}
\caption{Mean performance ratio (cheap / DIIM) by $k$ for Manpower (2\,000 trials per $k$).}
\label{tab:manpower-ratio-by-k}
\end{table}

\subsection{Close Rates}

Tables~\ref{tab:covid-close-by-k} and~\ref{tab:manpower-close-by-k} present the close rates (fraction of trials where $R_m \geq 0.80$).

\begin{table}[H]
\centering
\begin{tabular}{@{}lccccc@{}}
\toprule
\textbf{Method} & $k{=}3$ & $k{=}5$ & $k{=}7$ & $k{=}10$ & $k{=}12$ \\
\midrule
Total Output         & 66.1\% &  5.7\% & 62.0\% & 96.7\% & 100.0\% \\
PCA $\times$ xi      & 66.1\% & \textbf{64.1\%} & 62.0\% & 53.4\% &  99.8\% \\
PageRank $\times$ xi & 66.1\% &  5.7\% & 17.7\% & 96.7\% & 100.0\% \\
BL $\times$ xi       & 66.1\% &  5.7\% & \textbf{67.2\%} & 96.7\% & 100.0\% \\
FL $\times$ xi       & 66.1\% &  5.7\% & 62.0\% & 96.7\% &  99.8\% \\
\bottomrule
\end{tabular}
\caption{Close rates ($R_m \geq 0.80$) by $k$ for COVID-19 (2\,000 trials per $k$). Bold indicates notable structural method advantage.}
\label{tab:covid-close-by-k}
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{@{}lccccc@{}}
\toprule
\textbf{Method} & $k{=}3$ & $k{=}5$ & $k{=}7$ & $k{=}10$ & $k{=}12$ \\
\midrule
Total Output         & 16.4\% & 17.2\% & 97.1\% & 99.2\% & 100.0\% \\
PCA $\times$ xi      & 16.4\% & 17.2\% & 97.1\% & 99.2\% & 100.0\% \\
PageRank $\times$ xi & 16.4\% & 17.2\% & 33.6\% & 99.2\% & 100.0\% \\
BL $\times$ xi       & 16.4\% & 17.2\% & 97.1\% & 93.5\% & 100.0\% \\
FL $\times$ xi       & 16.4\% & 17.2\% & 97.1\% & 99.2\% & 100.0\% \\
\bottomrule
\end{tabular}
\caption{Close rates ($R_m \geq 0.80$) by $k$ for Manpower (2\,000 trials per $k$).}
\label{tab:manpower-close-by-k}
\end{table}


\subsection{Three Regimes of $k$}

The results reveal a striking non-monotonic pattern that can be decomposed into three regimes:

\begin{enumerate}
    \item \textbf{Small $k$ ($k = 3$)}: All cheap methods agree on the top 3 sectors and produce identical performance with a mean ratio of 0.870 (COVID-19) and 0.652 (Manpower). The dominant economic sectors (Manufacturing, Wholesale \& Retail, Transportation \& Storage) are so large that every method---whether based on size, linkage, or centrality---selects them. Two-thirds of trials achieve $\geq$80\% of DIIM in COVID-19 (66.1\%), but only 16.4\% in Manpower. Structural distinctions between cheap methods are irrelevant at this scale.

    \item \textbf{Medium $k$ ($k = 5$--$7$)}: The methods diverge, and structural information adds substantial value. At $k = 5$ in COVID-19, PCA $\times$ xi achieves a mean ratio of 0.817 versus 0.641 for Total Output---a 27\% relative improvement---and a close rate of 64.1\% versus just 5.7\%. This dramatic gap arises because PCA selects sector~3 (Construction) over sector~7 (Information \& Communications). At $k = 7$, BL $\times$ xi leads with a mean ratio of 0.834 and close rate of 67.2\%, capturing sector~12 (Public Administration) via its backward linkage signal. This is the regime where the ``marginal'' sectors (positions 4--7 in the ranking) matter most, and each method's structural lens produces meaningfully different selections.

    \item \textbf{Large $k$ ($k = 10$--$12$)}: The methods converge again as the intervention covers most of the economy. At $k = 12$ (80\% of 15 sectors), close rates reach 100\% for both scenarios, with mean ratios of 0.92--0.98. When intervening on nearly all sectors, the only question is which 3 sectors to \emph{exclude}, and even suboptimal exclusions have limited impact.
\end{enumerate}


\section{Key Observations}

Several patterns emerge from the multi-$k$ tables:

\begin{enumerate}
    \item \textbf{PCA $\times$ xi outperforms Total Output at $k = 5$ (COVID-19)}: PCA $\times$ xi achieves a mean ratio of 0.817 versus 0.641 for Total Output---a 27\% relative improvement---and a close rate of 64.1\% versus just 5.7\%. This stems from PCA's inclusion of sector~3 (Construction), which has the highest structural distance ($d_3^{\text{PCA}} = 0.789$) despite ranking only 6th in economic size. Construction's high initial inoperability ($q_{0,3} = 0.221$) makes intervening there disproportionately effective.

    \item \textbf{BL $\times$ xi leads at $k = 7$ (COVID-19)}: BL $\times$ xi achieves a mean ratio of 0.834 and a close rate of 67.2\%, outperforming all other cheap methods including PCA $\times$ xi (0.818, 62.0\%). Its backward linkage signal captures sector~12 (Public Administration), whose strong input dependencies amplify the cascading effect of intervention. This demonstrates that different structural methods have complementary strengths at different intervention scales.

    \item \textbf{PageRank $\times$ xi underperforms at medium $k$}: At $k = 7$, PageRank $\times$ xi drops to a mean ratio of 0.703 (COVID-19) and 0.749 (Manpower), with close rates of just 17.7\% and 33.6\% respectively---far below the other methods. PageRank's random-walk centrality prioritises sector~12 over sectors~3 and~10 at positions~6--7, which proves suboptimal under perturbed $q_0$.

    \item \textbf{All methods converge at large $k$}: At $k = 10$--$12$, close rates reach 97--100\% for most methods, with mean ratios of 0.91--0.98. The specific ranking matters little when the majority of sectors are covered.

    \item \textbf{No single cheap method dominates across all $k$}: PCA $\times$ xi is best at $k = 5$, BL $\times$ xi is best at $k = 7$, and Total Output matches the best at $k = 3$, $10$, and $12$. This suggests that the optimal cheap method depends on the intervention scale.
\end{enumerate}


\section{What Drives Cheap Method Performance?}

To understand \emph{why} cheap methods succeed or fail at a given $q_0$, we fit logistic regression models at $k = 5$ using features derived from the $q_0$ distribution:

\begin{equation}
    \log \frac{P(R_m \geq 0.80)}{1 - P(R_m \geq 0.80)} = \beta_0 + \beta_1 f_1 + \beta_2 f_2 + \cdots
    \label{eq:logistic-regression}
\end{equation}

\noindent where the features include the Gini coefficient, coefficient of variation (CV), and normalised entropy of $q_0$.

The analysis reveals scenario-specific patterns:

\begin{itemize}
    \item \textbf{COVID-19}: The concentration of $q_0$ is the key driver. When the disruption is concentrated in a few large sectors (high Gini), the cheap method's fixed ranking naturally aligns with the most impacted sectors. PCA $\times$ xi is less sensitive to this alignment because it identifies structurally important sectors (like Construction) regardless of $q_0$ concentration---hence its consistently high close rate of 64.1\% at $k = 5$.

    \item \textbf{Manpower}: The uniformity of $q_0$ matters most. Higher entropy (more evenly spread disruption) favours cheap methods because the ranking of intervention targets becomes less consequential when all sectors are similarly affected.
\end{itemize}

The practical implication is that cheap methods are most reliable when the disruption pattern aligns with the economic size distribution---a condition that is scenario-dependent and cannot be assumed in advance.


\section{Practical Implications}

The multi-$k$ analysis leads to the following practical guidelines:

\begin{enumerate}
    \item \textbf{Large-scale intervention ($k \geq 10$)}: Cheap methods are reliable substitutes for DIIM. With mean ratios above 0.90 and close rates of 97--100\%, the additional computational cost of DIIM is difficult to justify when intervening on the majority of sectors. Total Output alone---requiring no IO table analysis---achieves 95\% of DIIM's performance at $k = 12$.

    \item \textbf{Targeted intervention ($k = 3$--$7$)}: DIIM provides substantial value over purely size-based methods. At $k = 5$, Total Output achieves only 64\% of DIIM's loss reduction (close rate 5.7\%). However, structural methods significantly narrow this gap: PCA $\times$ xi achieves 82\% of DIIM (close rate 64.1\%), and BL $\times$ xi leads at $k = 7$ with 83\% of DIIM (close rate 67.2\%). While a gap to DIIM remains, structural methods provide a meaningful improvement over the naive size-based baseline in this regime.

    \item \textbf{Pre-event planning}: When $q_0$ is unknown, cheap methods remain valuable for pre-positioning resources. The Total Output ranking identifies the sectors whose disruption would cause the greatest absolute economic loss, while PCA $\times$ xi and BL $\times$ xi can flag structurally important sectors that size alone would miss (e.g., Construction in the COVID-19 scenario).

    \item \textbf{Crisis response}: Once $q_0$ is observed, run the full DIIM if targeting fewer than 10 sectors. The computational cost of the simulation is modest (seconds on modern hardware) and the performance gain is substantial.
\end{enumerate}

%=== END OF CHAPTER SIX ===
\newpage
