---
title: "DIIM Economic Loss Analysis — Functions"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    highlight: tango
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Overview

This document contains the core functions for the **Dynamic Inoperability Input–Output Model (DIIM)** analysis, including data loading utilities, the DIIM simulation engine, and comparison functions for ML-based vs. DIIM-based intervention strategies.

---

# 1. Load Libraries

```{r libraries}
library(openxlsx)
```

---

# 2. Data Loading Functions

## 2.1 `download_data` — COVID-19 Scenario (2019 IOT)

Loads the 2019 Input–Output Table, extracts the technical coefficient matrix **A**, total output **x**, final demand **c**, sector initial inoperability **q₀**, perturbation vector **c\***, and computes **A\***.

```{r download_data}
download_data <- function() {
    # download 2019 IOT and extract total output and final demand column
    data <- read.xlsx("dataset/Input_Output_Tables_2019_2021.xlsx", sheet = "2019IOTable")
    iot2019 <- data[2:(2 + 15 - 1), 3:(3 + 15 - 1)]
    x <- data[2:(2 + 15 - 1), ncol(data)]
    c <- data[2:(2 + 15 - 1), ncol(data) - 1]
    x <- as.matrix(sapply(x, as.numeric))
    c <- as.matrix(sapply(c, as.numeric))

    # extract technical coefficient matrix
    A <- read.xlsx("dataset/Input_Output_Tables_2019_2021.xlsx", sheet = "A")
    A <- A[2:(2 + 15 - 1), 3:(3 + 15 - 1)]
    A <- as.matrix(sapply(A, as.numeric))

    # extract sector initial inoperability
    q0_df <- read.xlsx("dataset/covid-data/Unemployment_and_Impact_Analysis.xlsx",
        sheet = "sector inoperability", colNames = FALSE
    )
    q0 <- as.numeric(q0_df[2:16, 3])

    c_star_df <- read.xlsx("dataset/covid-data/Unemployment_and_Impact_Analysis.xlsx",
        sheet = "c_star"
    )
    c_star <- as.numeric(c_star_df[1:15, 7])

    A_star <- solve(diag(as.vector(x))) %*% A %*% diag(as.vector(x))

    return(list(
        x = x,
        c = c,
        A = A,
        q0 = q0,
        c_star = c_star,
        A_star = A_star
    ))
}
```

## 2.2 `download_manpower_data` — Manpower Disruption Scenario (2022 IOT)

Loads the 2022 Input–Output Table and sector inoperability data for the manpower disruption analysis.

```{r download_manpower_data}
download_manpower_data <- function() {
    # Load 2022 IOT for A matrix and x vector
    data_A <- read.xlsx("dataset/manpower-disruption-data/2022_Input_Output_Table.xlsx",
        sheet = "A", colNames = FALSE
    )
    num_sectors <- nrow(data_A) - 1
    A <- data_A[2:(num_sectors + 1), 2:(num_sectors + 1)]
    A <- as.matrix(sapply(A, as.numeric))

    # Sheet "x" contains the total output
    data_x <- read.xlsx("dataset/manpower-disruption-data/2022_Input_Output_Table.xlsx",
        sheet = "x", colNames = FALSE
    )
    x <- data_x[2:(num_sectors + 1), 2]
    x <- as.numeric(x)

    # Load sector initial inoperability (q0)
    data_q0 <- read.xlsx("dataset/manpower-disruption-data/Sector_initial_inoperability.xlsx",
        sheet = "Sector_initial_inoperability", colNames = FALSE
    )
    q0 <- data_q0[2:(num_sectors + 1), 4]
    q0 <- as.numeric(q0)

    # Initialize c and c_star to 0 for manpower disruption scenario
    c <- rep(0, num_sectors)
    c_star <- rep(0, num_sectors)

    # Calculate A_star: A* = diag(x)^(-1) %*% A %*% diag(x)
    A_star <- solve(diag(x)) %*% A %*% diag(x)

    return(list(
        x = as.matrix(x),
        c = as.matrix(c),
        A = A,
        q0 = q0,
        c_star = c_star,
        A_star = A_star
    ))
}
```

---

# 3. DIIM Model

## 3.1 `DIIM` — Dynamic Inoperability Input–Output Model

Simulates sector inoperability evolution and cumulative economic loss over a specified time horizon. Supports optional intervention on key sectors (reducing their initial inoperability by 10%).

**Parameters:**

| Parameter | Description |
|---|---|
| `q0` | Initial inoperability vector |
| `A_star` | Interdependency matrix A* |
| `c_star` | External perturbation vector |
| `x` | Total output vector (yearly, in millions) |
| `lockdown_duration` | Number of days the external shock persists |
| `total_duration` | Total simulation horizon (days) |
| `key_sectors` | (Optional) Indices of sectors to reduce inoperability by 10% |
| `risk_management` | Scaling factor for recovery dynamics (default: 1) |
| `days_in_year` | Days in a year for daily output conversion (default: 366) |

```{r DIIM}
DIIM <- function(q0, A_star, c_star, x, lockdown_duration, total_duration,
                 key_sectors = NULL, risk_management = 1, days_in_year = 366) {
    a_ii <- diag(A_star)
    T <- total_duration

    num_sectors <- length(q0)
    inoperability_evolution <- matrix(NA, nrow = num_sectors, ncol = total_duration)

    # if key_sectors is not NULL, reduce those sectors' initial q0 by 10%
    if (!is.null(key_sectors)) {
        q0[key_sectors] <- q0[key_sectors] * 0.9
    }

    # we assume after T days the economic activity returns to 99% of pre-lockdown level
    q0[q0 == 0] <- 1e-8
    qT <- q0 * 1 / 100

    k <- log(q0 / qT) / (T * (1 - a_ii))
    K <- diag(as.vector(k))

    inoperability_evolution[, 1] <- q0

    for (t in 2:total_duration) {
        if (t <= lockdown_duration) {
            inoperability_evolution[, t] <- inoperability_evolution[, t - 1] +
                K %*% (A_star %*% inoperability_evolution[, t - 1] + c_star -
                    inoperability_evolution[, t - 1]) * risk_management
        } else {
            # after lockdown c* becomes 0 as no more external shock
            inoperability_evolution[, t] <- inoperability_evolution[, t - 1] +
                K %*% (A_star %*% inoperability_evolution[, t - 1] -
                    inoperability_evolution[, t - 1]) * risk_management
        }
    }

    # convert planned YEARLY output to DAILY output (in millions)
    x_daily <- x / days_in_year
    EL_evolution <- t(apply(inoperability_evolution, 1, cumsum)) * as.vector(x_daily)
    EL_end <- EL_evolution[, ncol(EL_evolution)]
    total_economic_loss <- sum(EL_evolution[, ncol(EL_evolution)])

    return(list(
        inoperability_evolution = inoperability_evolution,
        EL_evolution = EL_evolution,
        EL_end = EL_end,
        total_economic_loss = total_economic_loss
    ))
}
```

---

# 4. Simulation — ML vs. DIIM Comparison

## 4.1 `simulation_ml_vs_diim` — COVID-19 Scenario

Compares total economic loss under three strategies:

1. **No intervention** (baseline)
2. **DIIM-based intervention** — top 5 sectors by economic loss
3. **ML-based intervention** — key sectors identified via PCA (sectors 3, 2, 5, 8, 10)

```{r simulation_ml_vs_diim}
simulation_ml_vs_diim <- function(q0, A_star, c_star, x, lockdown_duration, total_duration) {
    # run model to calculate the economic loss (without intervention)
    model <- DIIM(q0, A_star, c_star, x, lockdown_duration, total_duration)
    EL_evolution <- model$EL_evolution

    # obtain the sectors with top economic loss
    max_econ_loss <- apply(EL_evolution, 1, max)
    sorted_indices <- order(max_econ_loss, decreasing = TRUE)
    top_econ_loss_5 <- sorted_indices[1:5]

    # rerun the model with intervention for the top economic sectors
    model_diim <- DIIM(q0, A_star, c_star, x, lockdown_duration, total_duration,
        key_sectors = top_econ_loss_5
    )

    ml_key_sectors <- c(3, 2, 5, 8, 10) # obtained from PCA
    model_ml <- DIIM(q0, A_star, c_star, x, lockdown_duration, total_duration,
        key_sectors = ml_key_sectors
    )

    model_tot_econ_loss <- model$total_economic_loss
    model_diim_tot_econ_loss <- model_diim$total_economic_loss
    model_ml_tot_econ_loss <- model_ml$total_economic_loss

    return(list(
        lockdown_duration = lockdown_duration,
        total_duration = total_duration,
        model_tot_econ_loss = model_tot_econ_loss,
        model_diim_tot_econ_loss = model_diim_tot_econ_loss,
        model_ml_tot_econ_loss = model_ml_tot_econ_loss
    ))
}
```

## 4.2 `simulation_ml_vs_diim_manpower` — Manpower Disruption Scenario

Same comparison as above but with ML key sectors identified for the manpower scenario (sectors 79, 107, 89, 11, 33).

```{r simulation_ml_vs_diim_manpower}
simulation_ml_vs_diim_manpower <- function(q0, A_star, c_star, x, lockdown_duration, total_duration) {
    # run model to calculate the economic loss (without intervention)
    model <- DIIM(q0, A_star, c_star, x, lockdown_duration, total_duration)
    EL_evolution <- model$EL_evolution

    # obtain the sectors with top economic loss
    max_econ_loss <- apply(EL_evolution, 1, max)
    sorted_indices <- order(max_econ_loss, decreasing = TRUE)
    top_econ_loss_5 <- sorted_indices[1:5]

    # rerun the model with intervention for the top economic sectors
    model_diim <- DIIM(q0, A_star, c_star, x, lockdown_duration, total_duration,
        key_sectors = top_econ_loss_5
    )

    ml_key_sectors <- c(79, 107, 89, 11, 33) # obtained from PCA
    model_ml <- DIIM(q0, A_star, c_star, x, lockdown_duration, total_duration,
        key_sectors = ml_key_sectors
    )

    model_tot_econ_loss <- model$total_economic_loss
    model_diim_tot_econ_loss <- model_diim$total_economic_loss
    model_ml_tot_econ_loss <- model_ml$total_economic_loss

    return(list(
        lockdown_duration = lockdown_duration,
        total_duration = total_duration,
        model_tot_econ_loss = model_tot_econ_loss,
        model_diim_tot_econ_loss = model_diim_tot_econ_loss,
        model_ml_tot_econ_loss = model_ml_tot_econ_loss
    ))
}
```
