%=== CHAPTER TWO (2) ===
%=== Methodology ===

\chapter{Methodology}

This chapter presents the analytical framework. Section~\ref{sec:diim-derivation} derives the DIIM from the Leontief input-output model. Section~\ref{sec:simplified-methods} introduces the simplified structural methods for key sector identification. Section~\ref{sec:performance-metric} defines the performance metric used to compare methods.

\section{The Dynamic Inoperability Input-Output Model}
\label{sec:diim-derivation}

\subsection{Inoperability}

This paper uses the DIIM model to assess the impacts of economic disruptions on the inoperability and economic losses of Singapore's sectors. The DIIM model is derived from the Leontief input-output model and is used to investigate the higher-order transmission effect of input-output linkages between sectors on inoperability. Using the DIIM model requires first introducing the concept of ``inoperability'', which is defined as follows:

\begin{equation}
    \label{eq:inoperability-definition}
    \text{Inoperability} = \frac{\text{As Planned Production} - \text{Degraded Production}}{\text{As Planned Production}}
\end{equation}

Where ``As Planned Production'' represents the output level of a sector under normal production, and ``Degraded Production'' represents the output level of a sector after being shocked in its production process. ``Inoperability'' takes values between 0 and 1, with higher values indicating greater damage to production caused by the shock. A value of 1 means the shocked sector has completely lost its production capacity, while a value of 0 means the sector is producing at a normal level.

\subsection{From Leontief to IIM}

Using the concept of inoperability defined above, the IIM model can be derived from the Leontief input-output model \cite{Leontief1936}, as shown in Equation~\eqref{eq:leontief-model}.

\begin{equation}
    x = A x + c
    \label{eq:leontief-model}
\end{equation}

Where $x$ is the total output vector; $A$ is the technical coefficient matrix; $c$ is the final demand vector. If we define the output levels and the final demand vector of some shocked sectors as $\tilde{x}$ and $\tilde{c}$ respectively, then we can construct the input-output relationship between sectors after the shock.

Subtracting Equation~\eqref{eq:leontief-model} from the shocked-economy formulation:

\begin{equation}
    x - \tilde{x} = A (x - \tilde{x}) + (c - \tilde{c})
    \label{eq:leontief-difference}
\end{equation}

Defining $\hat{x}$ as a diagonalized matrix of the output vector and left multiplying $\hat{x}^{-1}$ on both sides of Equation~\eqref{eq:leontief-difference}:

\begin{equation}
    \hat{x}^{-1}(x - \tilde{x}) = \hat{x}^{-1}A(x - \tilde{x}) + \hat{x}^{-1}(c - \tilde{c})
    \label{eq:normalized-difference}
\end{equation}

Let $q = \hat{x}^{-1}\left(x - \tilde{x}\right), \quad A^{*} = \hat{x}^{-1} A \hat{x}, \quad c^{*} = \hat{x}^{-1}\left(c - \tilde{c}\right)$; then Equation~\eqref{eq:iim-model} can be derived, which is the IIM model.

\begin{equation}
    q = A^{*} q + c^{*}
    \label{eq:iim-model}
\end{equation}

The details of each matrix in the IIM model are shown in Equations~\eqref{eq:q-definition}, \eqref{eq:astar-definition}, and~\eqref{eq:cstar-definition}:

\begin{equation}
    \label{eq:q-definition}
    q = \hat{x}^{-1} (x - \tilde{x}) =
    \begin{bmatrix}
        \dfrac{1}{x_1} & 0 & \cdots & \cdots & 0 \\
        0 & \ddots & & & \vdots \\
        \vdots & & \dfrac{1}{x_i} & & \vdots \\
        \vdots & & & \ddots & 0 \\
        0 & \cdots & \cdots & 0 & \dfrac{1}{x_n}
    \end{bmatrix}
    \begin{bmatrix}
        x_1 - \tilde{x}_1 \\
        \vdots \\
        x_i - \tilde{x}_i \\
        \vdots \\
        x_n - \tilde{x}_n
    \end{bmatrix}
\end{equation}

\begin{equation}
    A^{*} = \hat{x}^{-1} A \hat{x} =
    \begin{bmatrix}
        a_{11}\dfrac{x_{1}}{x_{1}} & \cdots & a_{1j}\dfrac{x_{j}}{x_{1}} & \cdots & a_{1n}\dfrac{x_{n}}{x_{1}} \\
        \vdots & \ddots & \vdots & \ddots & \vdots \\
        a_{i1}\dfrac{x_{1}}{x_{i}} & \cdots & a_{ij}\dfrac{x_{j}}{x_{i}} & \cdots & a_{in}\dfrac{x_{n}}{x_{i}} \\
        \vdots & \ddots & \vdots & \ddots & \vdots \\
        a_{n1}\dfrac{x_{1}}{x_{n}} & \cdots & a_{nj}\dfrac{x_{j}}{x_{n}} & \cdots & a_{nn}\dfrac{x_{n}}{x_{n}}
    \end{bmatrix}
    \label{eq:astar-definition}
\end{equation}


\begin{equation}
    \label{eq:cstar-definition}
    \mathbf{c}^{*} = \hat{\mathbf{x}}^{-1} \left( \mathbf{c} - \tilde{\mathbf{c}} \right)
    =
    \begin{bmatrix}
        \dfrac{1}{x_{1}} & 0 & \cdots & 0 \\
        0 & \ddots & \ddots & \vdots \\
        \vdots & \ddots & \dfrac{1}{x_{i}} & 0 \\
        \vdots & & \ddots & \vdots \\
        0 & \cdots & \cdots & \dfrac{1}{x_{n}}
    \end{bmatrix}
    \begin{bmatrix}
        c_{1} - \tilde{c}_{1} \\
        \vdots \\
        c_{i} - \tilde{c}_{i} \\
        \vdots \\
        c_{n} - \tilde{c}_{n}
    \end{bmatrix}
\end{equation}

\subsection{The DIIM}

The IIM model can be extended to the DIIM model by introducing a time variable and an elasticity coefficient matrix describing the recovery capacity of sectors \cite{SantosHaimes2004}. The discrete form of the DIIM model is:

\begin{equation}
    q(t+1) = q(t) + K \big[ A^{*} q(t) + c^{*}(t) - q(t) \big]
    \label{eq:diim-discrete}
\end{equation}

Where $K$ is the elasticity coefficient matrix describing the recovery capacity of sectors after being shocked in production. Assuming that the recovery capacity of a sector depends only on its own production, and is unrelated to the production linkages with other sectors, thus $K$ is a diagonal matrix with diagonal elements greater than 0. The larger the diagonal elements, the stronger the recovery capacity of the corresponding sector in response to shocks. $t$ is the discrete time variable, and $q(t)$ represents the sectoral inoperability vector at time~$t$. Approximating Equation~\eqref{eq:diim-discrete} to differential form:

\begin{equation}
    \label{eq:diim-continuous}
    \dot{q}(t) = K \bigl[ A^{*} q(t) + c^{*}(t) - q(t) \bigr]
\end{equation}

Solving Equation~\eqref{eq:diim-continuous} we can get the equation describing the evolution of sectoral inoperability over time:

\begin{equation}
    q(t) = e^{-K(I-A^{*})t} q(0) + \int_{0}^{t} K e^{-K(I-A^{*})(t-z)} c^{*}(z)\, dz
    \label{eq:diim-solution-general}
\end{equation}

Assuming that the demand shock $c^{*}$ remains unchanged, and $c^{*}=0$, then Equation~\eqref{eq:diim-solution-general} can be simplified to:

\begin{equation}
    q(t) = e^{-K (I - A^{*}) t} q(0)
    \label{eq:diim-solution-homogeneous}
\end{equation}

Where $q(0)$ represents the initial inoperability vector of sectors after being shocked. As time goes by, the inoperability changes at a rate of $e^{-K(I - A^{*})t}$.

\subsection{Initial Inoperability}

The initial inoperability $q(0)$ is calculated as:

\begin{equation}
    \label{eq:initial-inoperability}
    q_i(0) = \underbrace{\left(1 - \text{RLI}_i\right)\left(1 - \text{ESS}_i\right)}_{\text{Workforce Unavailability}} \times \frac{\text{LAPI}_i}{\text{Sector Output}_i}
\end{equation}

The first factor, workforce unavailability, is adapted from \cite{DelRioChanona2020,Pichler2022}. $\text{RLI}_i$ is the Remote Labor Index, representing the probability that a worker in sector~$i$ can work from home. $\text{ESS}_i$ is the Essential Score, representing the probability that a worker has an essential job that permits on-site work during a lockdown. The product $(1-\text{RLI}_i)(1-\text{ESS}_i)$ gives the expected fraction of workers who cannot work during the lockdown.

The second factor, $\text{LAPI}_i / \text{Sector Output}_i$, captures how dependent each sector is on labour as an input, where LAPI is the Local Area Personal Income (equivalent to compensation of employees in the IO table).

\subsection{Elasticity Coefficient}

According to Equation~\eqref{eq:diim-solution-homogeneous}, the inoperability of sector~$i$ at time~$t$ is:

\begin{equation}
    q_i(t) = q_i(0) \, e^{-k_i \left( 1 - a_{ii}^{*} \right) t}
    \label{eq:qi-exponential-decay}
\end{equation}

From Equation~\eqref{eq:qi-exponential-decay}, we can derive the calculation formula for~$k_i$:

\begin{equation}
    k_i = \frac{\ln\left[q_i(0)/q_i(T)\right]}{T_i\left(1 - a_{ii}^{*}\right)}
    \label{eq:ki-definition}
\end{equation}

Where $q_i(0)$ is the initial inoperability of sector~$i$; $q_i(T)$ is the inoperability of sector~$i$ at time~$T$; $a_{ii}^{*}$ is the element on the diagonal of matrix~$A^{*}$, representing the self-dependence of sector~$i$.

\subsection{Economic Loss}

Under the analytical framework of the DIIM model, the cumulative economic loss for sector~$i$ is:

\begin{equation}
    EL_i = x_i \int_{t=0}^{t=T} q_i(t) \, dt
    \label{eq:economic-loss}
\end{equation}

Where $x_i$ is the planned output of sector~$i$ under normal production.


\section{Key Sector Identification Methods}
\label{sec:simplified-methods}

A key sector is defined as one where targeted intervention---modelled as a 10\% reduction in initial inoperability $q(0)$---produces the largest reduction in total economic loss. Below we first describe the intervention and loss reduction evaluation procedure, followed by the expensive benchmark and five simplified structural alternatives.

\subsection{Intervention and Loss Reduction Evaluation}
\label{sec:intervention-procedure}

Each method produces a ranking of sectors. To evaluate a method's effectiveness, we select its top-$k$ ranked sectors and simulate a targeted intervention: the initial inoperability of each selected sector is reduced by 10\%. The DIIM is then re-run with this modified $q(0)$ to obtain the post-intervention total economic loss. The difference between the baseline loss (no intervention) and the post-intervention loss gives the economic loss reduction achieved by that method's top-$k$ selection. This procedure is applied identically to the DIIM benchmark and to every simplified method, so that all methods are compared on the same basis.

The formal procedure is given in Algorithm~\ref{alg:loss-reduction}.

\begin{algorithm}[H]
\caption{Intervention and Loss Reduction Evaluation}\label{alg:loss-reduction}
\KwIn{Initial inoperability $q(0)$, interdependency matrix $A^*$, demand perturbation $c^*$, total output vector $x$, number of key sectors $k$, sector ranking method $m$}
\KwOut{Percentage loss reduction $\%\Delta EL_m$ for method $m$}
\BlankLine
\tcp{Step 1: Compute baseline economic loss (no intervention)}
$\text{model}_{\text{base}} \leftarrow \text{\textsc{DIIM}}\bigl(q(0),\, A^*,\, c^*,\, x\bigr)$\;
$EL_{\text{base}} \leftarrow \text{TotalEconomicLoss}\bigl(\text{model}_{\text{base}}\bigr)$\;
\BlankLine
\tcp{Step 2: Rank sectors using method $m$ and select top-$k$}
$\text{ranking} \leftarrow \text{\textsc{RankSectors}}(m)$\;
$S_k \leftarrow \text{ranking}[1:k]$ \tcp*{top-$k$ sector indices}
\BlankLine
\tcp{Step 3: Apply intervention (reduce $q(0)$ by 10\% for selected sectors)}
$q'(0) \leftarrow q(0)$\;
\For{each sector $i \in S_k$}{
  $q'_i(0) \leftarrow 0.9 \times q_i(0)$\;
}
\BlankLine
\tcp{Step 4: Re-run DIIM with the modified initial inoperability}
$\text{model}_{\text{int}} \leftarrow \text{\textsc{DIIM}}\bigl(q'(0),\, A^*,\, c^*,\, x\bigr)$\;
$EL_{\text{int}} \leftarrow \text{TotalEconomicLoss}\bigl(\text{model}_{\text{int}}\bigr)$\;
\BlankLine
\tcp{Step 5: Compute loss reduction}
$\Delta EL_m \leftarrow EL_{\text{base}} - EL_{\text{int}}$\;
$\%\Delta EL_m \leftarrow \dfrac{\Delta EL_m}{EL_{\text{base}}} \times 100$\;
\BlankLine
\Return $\%\Delta EL_m$\;
\end{algorithm}

In words, the algorithm proceeds as follows. First, the DIIM is run with the original $q(0)$ to establish the baseline total economic loss, which represents the scenario with no policy intervention. Second, a sector ranking method $m$ (either DIIM or a simplified method) produces an ordered list of sectors, and the top $k$ are selected. Third, the initial inoperability of each selected sector is reduced by 10\%, representing a targeted policy intervention that partially restores the sector's production capacity at time zero. Fourth, the DIIM is re-run with this modified initial condition $q'(0)$ to obtain the post-intervention total economic loss. Finally, the loss reduction is computed as the difference between the baseline and post-intervention losses. This procedure is repeated for every method $m$, with identical $q(0)$, $k$, and intervention magnitude, so that differences in loss reduction reflect only the quality of each method's sector ranking.

\subsection{DIIM-Based Identification (Benchmark)}

The DIIM identifies key sectors by running the full simulation without intervention, then ranking sectors by their cumulative economic loss $EL_i$ from Equation~\eqref{eq:economic-loss}. The top-$k$ sectors are selected for intervention. This is the ``gold standard'' but requires knowledge of $q(0)$ and~$c^{*}$.

\subsection{Total Output ($x_i$)}

This is the simplest benchmark method. Sectors are ranked simply by their total planned economic output $x_i$. This method ignores the network structure entirely and assumes that intervening in the largest economic sectors will yield the greatest reduction in loss.

\subsection{PCA $\times\, x_i$}

Instead of using $A$ directly, this approach constructs the integrated input-output coefficient matrix:

\begin{equation}
    H = A (I - A)^{-1} = A \cdot L
    \label{eq:integrated-io-matrix}
\end{equation}

where $L = (I - A)^{-1}$ is the Leontief inverse. The matrix $H$ captures both direct and indirect effects through the production network. PCA is applied to~$H$, and a structural distance is computed for each sector based on its Euclidean distance from the origin in the space of the principal components:

\begin{equation}
    d_i^{\text{PCA}} = \sqrt{ \sum_{k=1}^{n\_pcs} v_{ik}^2 }
    \label{eq:pca-distance}
\end{equation}

where $v_{ik}$ is the loading of sector~$i$ on principal component $k$. To account for economic size, this structural distance is weighted by the sector's total output $x_i$:

\begin{equation}
    \text{Score}_i^{\text{PCA}} = d_i^{\text{PCA}} \times x_i
    \label{eq:pca-xi-score}
\end{equation}

Sectors with larger $\text{Score}_i^{\text{PCA}}$ are ranked higher.

\subsection{PageRank $\times\, x_i$}

The interdependency matrix $A^*$ is treated as the weighted adjacency matrix of a directed graph. The PageRank algorithm \cite{BrinPage1998} is applied to compute a structural centrality score $PR_i$ for each sector, which measures importance based on the structure of incoming links with a damping factor. This centrality score is then weighted by the sector's total output:

\begin{equation}
    \text{Score}_i^{\text{PR}} = PR_i \times x_i
    \label{eq:pagerank-xi}
\end{equation}

\subsection{Backward Linkage $\times\, x_i$ (BL $\times\, x_i$)}

Backward linkages measure the extent to which a sector draws on other sectors' outputs as inputs \cite{Rasmussen1956,Chenery1958}. They are calculated as the column sums of the Leontief inverse:

\begin{equation}
    BL_j = \sum_{i=1}^{n} l_{ij}
    \label{eq:backward-linkage}
\end{equation}

where $l_{ij}$ are elements of $L$. A high $BL_j$ indicates that sector~$j$ generates large demand for inputs from other sectors. To identify key sectors, this structural linkage is weighted by total output:

\begin{equation}
    \text{Score}_j^{\text{BL}} = BL_j \times x_j
\end{equation}

\subsection{Forward Linkage $\times\, x_i$ (FL $\times\, x_i$)}

Forward linkages measure how much a sector's output is used as inputs by other sectors \cite{Rasmussen1956,Dietzenbacher2002}. They are calculated as the row sums of the Leontief inverse:

\begin{equation}
    FL_i = \sum_{j=1}^{n} l_{ij}
    \label{eq:forward-linkage}
\end{equation}

A high $FL_i$ indicates that sector~$i$ is an important supplier to the rest of the economy. This structural linkage is weighted by total output:

\begin{equation}
    \text{Score}_i^{\text{FL}} = FL_i \times x_i
\end{equation}

\section{Performance Metric}
\label{sec:performance-metric}

To compare a simplified method against DIIM, we define the \emph{performance ratio}:

\begin{equation}
    R_m = \frac{\Delta EL_m}{\Delta EL_{\text{DIIM}}}
    \label{eq:performance-ratio}
\end{equation}

where $\Delta EL_m$ is the total economic loss reduction achieved by intervening in the top-$k$ sectors identified by simplified method~$m$, and $\Delta EL_{\text{DIIM}}$ is the reduction achieved by the DIIM's top-$k$ sectors. If $R_m \geq 0.95$, the simplified method is considered ``close enough'' to substitute for DIIM.

%=== END OF CHAPTER TWO ===
\newpage
